Automatically Splitting a Two-Stage Lambda Calculus

### 2015-07-11 paper/relatedwork.tex
similar to --> much like

In an attempt to reduce the complexity of the specialization phase of
self-applicable partial evaluators, researchers explored the use of
staging transformations that, similar to splitting, partition an input
program into two programs: one that produces static outputs from
static inputs, and a second that produces dynamic outputs from both
static and dynamic inputs.  These metastatic transformations, referred
to as \emph{program bifurcation}\,\cite{DeNiel91} or as \emph{binding
  time separation}\,\cite{Mogensen89a}), serve as a preprocessing step
to specialization and were successfully applied to functional
languages that supported recursion\,\cite{DeNiel91}. However, unlike
splitting, where the goal is to emit code where first stage results
are computed once and then reused in multiple invocations of second
stage execution, the second function produced by program bifurcation
does not use the results of the first; Instead,
%only serves as a preprocessing step for the subsequent
%specialization phase of partial evaluation.
it has access to the first stage inputs and recomputes all required
first stage computations. (As noted by Knoblock and
Ruf\,\cite{knoblock96}, it may be possible to modify the program
bifurcation algorithm to cache and reuse the intermediate results, but
this was never attempted.) The grammar-based binding-time
specifications used in program bifurcation were capable of describing
data structures with purely-static, purely-dynamic, and mixed stage
content, much like the type system of \lang.

>>>>

In an attempt to reduce the complexity of the specialization phase of
self-applicable partial evaluators, researchers explored the use of
staging transformations that, much like splitting, partition an input
program into two programs: one that produces static outputs from
static inputs, and a second that produces dynamic outputs from both
static and dynamic inputs.  These metastatic transformations, referred
to as \emph{program bifurcation}\,\cite{DeNiel91} or as \emph{binding
  time separation}\,\cite{Mogensen89a}), serve as a preprocessing step
to specialization and were successfully applied to functional
languages that supported recursion\,\cite{DeNiel91}. However, unlike
splitting, where the goal is to emit code where first stage results
are computed once and then reused in multiple invocations of second
stage execution, the second function produced by program bifurcation
does not use the results of the first; Instead,
%only serves as a preprocessing step for the subsequent
%specialization phase of partial evaluation.
it has access to the first stage inputs and recomputes all required
first stage computations. (As noted by Knoblock and
Ruf\,\cite{knoblock96}, it may be possible to modify the program
bifurcation algorithm to cache and reuse the intermediate results, but
this was never attempted.) The grammar-based binding-time
specifications used in program bifurcation were capable of describing
data structures with purely-static, purely-dynamic, and mixed stage
content, much like the type system of \lang.
